---
title: "USL Part 2: PCA"
author: "Evrim Bilgen"
date: "12/17/2021"
output: html_document
---
<style>
#TOC
{
color:#00008B;
font-family: Calibri;
font-size: 16px;
border-color: #708090;
}

h1.title {
  color: #000000;
  background-color: #F5F5F5;
  opacity: 0.6;
  font-family: Calibri;
  font-size: 25px;
}
h4.author{
color:#000000;
font-family: Calibri;
background-color: #F5F5F5;
}
body {
color: #000000;
font-family: Calibri;
background-color: #F5F5F5;
}
  pre{
     color: #000000;
      background-color: #F8F8FF;
     }
</style>

# **PCA ANALYSIS** 

## Our packages

```{r }

library(readr)
library(dplyr)
library(ggplot2)
library(corrplot)
library(gridExtra)
library(pROC)
library(MASS)
library(caTools)
library(caret)
library(caretEnsemble)
library(doMC)
library(DataExplorer)



```


## Our Data & Cleaning & Manipulating (EDA)
```{r }

cardio = read.csv("/Users/ewrimmm/Desktop/R exercises/USL_21/cardio.csv") 



```

```{r include=FALSE}
cardio <- cardio[, c('sex','target','slope','exang','restecg','fbs','cp','ca','thal','age','trestbps','chol','thalach','oldpeak')]
```


```{r }
attach(cardio)

glimpse(cardio)

```


Now, looking at out data's structure

```{r }
str(cardio)
```

Checking the missing values

```{r }
plot_missing(cardio)

cat("There are", sum(is.na(cardio)), "N/A values.")

```
There aren’t NA values.



Now, it's time to change the sex and target columns for easy interpretation

```{r }

cardio$sex <- factor(cardio$sex, levels = c(0,1), labels = c("Female","Male"))



```



```{r }

cardio_2 <-cardio

cardio_2$target <- factor(cardio_2$target, levels = c(0,1), labels = c("No","Yes"))

```


In a nutshell, our data:
```{r }
summary(cardio)

```


Target variable indicates that our candidate is a target for disease or not. So lets look at this variable


```{r }
prop.table(table(cardio$target))

```
seems almost well balanced



#Data Visualization

```{r }
corr_mat <- cor(cardio[,7:ncol(cardio)])

corrplot(corr_mat, order = "hclust", tl.cex = 1, addrect = 8)

```

As we can see here, there is good correlation between some variables.



Lets try another plotting type for our numeric variables to see outliers

```{r }
boxplot(cardio[,3:13])
```


And now, we can visualize the prediction ratio.

```{r }
ggplot(cardio, aes(x = target, fill = target)) +
  geom_bar(position = 'dodge', color = 'black') +
  ggtitle('Prediction analysis in statistical sample') +
  xlab('Possibility of Cardiovasculer Disease') +
  ylab('Total') +
  theme_minimal() +
  theme(legend.position = 'right') +
  guides(fill = guide_legend(nrow = 2,byrow = TRUE)) +
  scale_fill_discrete(name = " Cardiovasculer Diseases", labels = c("No", "Yes"))


```


See the above analysis in another point of view:

```{r }
ggplot(cardio, aes(x=cardio$target, fill=cardio$sex)) + 
  geom_bar() +
  xlab("Cardiovasculer Disease") +
  ylab("Count") +
  ggtitle("Analysis of possibility of cardiovasculer disease") +
  scale_fill_discrete(name = "Heart Disease", labels = c("No", "Yes"))

```


Disturbution of female and male sex analysis. 

```{r }

x=table(cardio$sex)
par(mfrow=c(1,2))
barplot(x,
        col=rainbow(2),
        legend=rownames(x),
        xlab="Female and Male",
        ylab="Count",
        main="Sex ")

pct=round(x/sum(x)*100)
lbs=paste(c("Female","Male")," ",pct,"%",sep=" ")
pie(x,labels=lbs,main="Percentage of Male and Female")


```


```{r }

range(cardio$age)

hist(cardio$age,labels=TRUE,main=" Age Distrubution",xlab="Age Groups",ylab="Frequency",col="purple")


```


So here we can see male who are between 45-55 is more likely to get cardiovasculer disease. So males are tend to be get cardiovasculer diseases more then female

```{r }

cardio %>%
  ggplot(aes(x=age,fill=sex)) +
  geom_histogram(bins = 20) + 
  xlab("Age") + 
  ylab("Count")+
  guides(fill = guide_legend(title = "Sex"))


```


Lets see more deeper analysis in some feature related to cardiovasuler diseases

```{r }
p1 <- cardio %>%
  ggplot(aes(x=age,y=chol,color=sex, size=chol))+
  geom_point(alpha=0.5)+xlab("Age") +
  ylab("Cholesterol")+
  ggtitle('Cholesterol in mg/dl in sex distribution') +
  guides(fill = guide_legend(title = "Sex"))

p2 <-cardio %>%
  ggplot(aes(x=age,y=thalach,color=sex, size=thalach))+
  geom_point(alpha=0.5)+xlab("Age") +
  xlab("Age") +
  ylab("Max. Heart Rate in sex distribution")+
  guides(fill = guide_legend(title = "Sex"))



p3 <-cardio %>%
  ggplot(aes(x=age,y=cp,color=sex, size=cp))+
  geom_point(alpha=0.5)+xlab("Age") +
  xlab("Age") +
  ylab("Chest Pain in sex distribution")+
  guides(fill = guide_legend(title = "Sex"))


p4 <-cardio %>%
  ggplot(aes(x=age,y=exang,color=sex, size=exang))+
  geom_point(alpha=0.5)+xlab("Age") +
  xlab("Age") +
  ylab("exercise induced angina in sex distribution")+
  guides(fill = guide_legend(title = "Sex"))
```


```{r }
grid.arrange(p1,p2, ncol=2)
```

```{r }
grid.arrange(p3,p4,ncol=2)
```



#PCA - PRINCIPLE COMPONENT ANALYSIS


While our dependent variable is Target which has 2 values : 0 and 1. 
  * 0 means there is no heart attack 
  * 1 means there is Heart Attack.
  
  
  We are going to use this variable for our PCA analysis
  
  
```{r }
set.seed(1234)
data_i <- createDataPartition(cardio$target, p=0.7, list = FALSE)
traindata <- cardio[data_i, -1]
testdata <- cardio[-data_i, -1]

```

```{r }
pca_res <- prcomp(cardio[,3:ncol(cardio)], center = TRUE, scale = TRUE)
plot(pca_res, type="l")

```

Lets see the results
```{r }
summary(pca_res)
```

We are going to pick first two  components which explains the 0.3514 of the variance. We need almost 11 components to explain around %96 of the variance.

```{r }
pca_data <- as.data.frame(pca_res$x)

ggplot(pca_data, aes(x=PC1, y=PC2, col=cardio_2$target)) + geom_point(alpha=0.5) + labs(colour = "Is she/he a target?")

```


We can see here the density of the two components

```{r }

target <- cardio_2$target

pc1 <- ggplot(pca_data, aes(x=PC1, fill=target)) + geom_density(alpha=0.5)  + ggtitle('Density in PC1')
              
pc2 <- ggplot(pca_data, aes(x=PC2, fill=target)) + geom_density(alpha=0.5)  + ggtitle('Density in PC2')

grid.arrange(pc1, pc2, ncol=2)

```
## Referances

 * https://www.datacamp.com/
 * Principal Component Analysis in R Youtube by Hefin Rhys
 * kaggle
 * https://stackoverflow.com/


