---
title: 'USL Project/ Clustering-PCA-Market Analysis/ Part 1: Clustering'
author: "Evrim Bilgen"
date: "12/16/2021"
output: html_document
---

<style>
#TOC
{
color:#00008B;
font-family: Calibri;
font-size: 16px;
border-color: #708090;
}

h1.title {
  color: #000000;
  background-color: #F5F5F5;
  opacity: 0.6;
  font-family: Calibri;
  font-size: 25px;
}
h4.author{
color:#000000;
font-family: Calibri;
background-color: #F5F5F5;
}
body {
color: #000000;
font-family: Calibri;
background-color: #F5F5F5;
}
  pre{
     color: #000000;
      background-color: #F8F8FF;
     }
</style>

# **UNSUPERVISED LEARNING** 

## **Content**{-}

This notebook contains these subjects;


* Unsupervised Learning with Clustering
* Dimensionality reduction &Principal component analysis (PCA)
* Market basket analysis 

## **Introduction to dataset**

In this section, I will demonstrate an unsupervised learning analysis using xxx from yyy. The analysis includes clustering using K-means algorithm 


## **Data Analysis** 
### Import Packages

```{r }

library(tidyverse)
library(dplyr)
library(ggExtra)
library(DT)
library(cluster)
library(gridExtra)
library(grid)
library(plotly)
library(corrgram)
library(wordcloud)
library(wordcloud2)
library(tm)
library(tmap)
library(magrittr)
library(data.table)
library(corpus)
library(corrplot)
library(ggplot2); theme_set(theme_bw())
library(skmeans)
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(GGally)
library(plotly)


```

### Read Data

```{r }

     df = read.csv("/Users/ewrimmm/Desktop/R exercises/USL_21/IMDB.csv") 

     attach(df)

     summary(df)

     glimpse(df)

```

Clearing N/A Values:

```{r }
cat("There are", sum(is.na(df)), "N/A values.")

df <- df %>% mutate(across(everything(), .fns = ~replace_na(.,0))) 

```

Check out the result:
```{r }
df%>%summarise_all(funs(n_distinct))
```

Also we look at the structure of the data:

```{r }
str(df)
```
Some of variables are characters and others are numerical.


### Data Wrangling & Cleaning $ Fixing
```{r }
colnames(df)

unique(df$name)

unique(df$rating)

df <- df[!df$rating %in% c("Not Rated", "Unrated"), ]


```

### Data & Cluster Analysis

Visualization of variables

```{r }
df %>%
  keep(is.numeric) %>%                     
  gather() %>%                            
  ggplot(aes(value)) +                     
  geom_histogram(fill="green", colour="black") +
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_density() +
  labs(x="Values", y="Frequency")


```

Trying to visualize the  numeric variables:

```{r }
df %>%
  keep(is.numeric) %>%                     
  gather() %>%                            
  ggplot(aes(value)) +                     
  geom_histogram(fill="green", colour="black") +
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_density() +
  labs(x="Values", y="Frequency")
```

```{r }
df_numericVars <- which(sapply(df, is.numeric))
df_numericVarNames <- names(df_numericVars)
```

Then we have
```{r }
df_numericVars <- which(sapply(df, is.numeric))
df_numericVarNames <- names(df_numericVars)

paste("There are ", length(df_numericVars) ," numeric variable in datasets")

print(df_numericVarNames)

```

Lets plot them all!
```{r }
ggplot(df, aes(genre, fill = genre)) + geom_bar() + theme_classic()

```

Interesting! Action, comedy and drama genres are leading the leauge.

See the whole picture:
```{r }
df %>% 
  ggplot(aes(score, fill = genre)) +
  geom_histogram(aes(y=..density..), alpha= .35, 
                 position="identity", show.legend = F) +
  geom_density(aes(colour = genre), 
               stat = "density", alpha = .3,
               show.legend = F) +
  facet_wrap(~genre, scales = "free")

```

Lets dig deeper to find a relationship:

```{r, eval = F}
th1<- theme_bw()+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
                       legend.position="top")


th2<- theme_bw()+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
                       legend.position="none")


plot_grid(ggplot(df, aes(budget, gross)) + 
            geom_point(alpha = 0.4) + geom_smooth() + scale_x_log10()+th1,
          ggplot(df, aes(gross, runtime)) + 
            geom_point(alpha = 0.4) + geom_smooth() + scale_x_log10()+th1,
          ggplot(df, aes(gross, score)) + 
            geom_point(alpha = 0.4) + geom_smooth() + scale_x_log10()+th1,
          ggplot(df, aes(budget, score)) + 
            geom_point(alpha = 0.4) + geom_smooth() + scale_x_log10()+th2,
          ggplot(df, aes(budget, runtime)) + 
            geom_point(alpha = 0.4) + geom_smooth() + scale_x_log10()+th2,
          align = "h" )


```

The most powerful plot is the first one which means the higher the budget, the higher the amount grossed.
It also followed by gross-score couple. So we can use these two variables for k-means later on.


Can The score variable would be a good indicator of the targeted genres?

```{r }
  summary(score)
ggplot(as.data.frame(score), aes(y = score)) + geom_boxplot(fill='#F8766D')

```


```{r }
ggplot(df, aes( x = score, fill = genre)) + geom_density(alpha = 0.4)

```
It is quite interesting that seeing the music genre is strong among others.


Go back to our first related couple(gross & budget)
```{r }

rel_1 <- df %>% group_by(year) %>% 
  summarise(avg_budget = mean(budget, na.rm = TRUE),
            avg_gross = mean(gross, na.rm = TRUE)) %>%
  pivot_longer(avg_budget:avg_gross, names_to = "class") 

rel_1$class2 <- factor(rel_1$class, levels = c("avg_gross", "avg_budget"))

rel_1 %>% ggplot() + 
  geom_area(aes(year, value, fill = class2)) + 
  scale_fill_discrete(name = "", labels = c("Gross", "Budget"))

rm(rel_1)
```

So we can be sure that throughout the years, both are increasing but mostly gross is dominating


```{r }

p1 <- ggplot(as.data.frame(gross), aes(y = gross)) + geom_boxplot(fill='#F8766D') + ylim(c(1,1500))
p2 <- ggplot(as.data.frame(budget), aes(y = budget)) + geom_boxplot(fill='#00BFC4') + ylim(c(1,150000))
grid.arrange(p1, p2, ncol = 2)
```


### K-Means

My hypothesis is here finding and  answer that using by k-means is that there is an bond that movies can be grouped (clustered) according to their budget given their gross. 
So the null hypothesis of this will be the opposite which means no clusters between these features

```{r }

Kdata <- df[,c(12,13)]

```

using Elbow approach for K estimation

```{r } 

tot.withinss <- vector("numeric", length = 10)
for (i in 1:10){
  kDet <- kmeans(Kdata, i)
  tot.withinss[i] <- kDet$tot.withinss
}

```




```{r } 

ggplot(as.data.frame(tot.withinss), aes(x = seq(1,10), y = tot.withinss)) + 
  geom_point(col = "#F8766D") +    
  geom_line(col = "#F8766D") + 
  theme(axis.title.x.bottom = element_blank()) +
  ylab("Within-cluster Sum of Squares") +
  xlab("Number of Clusters") +
  ggtitle("Elbow K Estimation")
```
Seeing the plot, there is reasonable point for the K value which would be k=5



So we are going to cluster the data by 5

```{r }
df_Clusters <- kmeans(Kdata, 5)
df_Clusters
```

Then visualize the clusters
```{r }

ggplot(Kdata, aes(x = gross, y = budget)) + 
  geom_point(stat = "identity", aes(color = as.factor(df_Clusters$cluster))) +
  scale_color_discrete(name=" ",
                       breaks=c("1", "2", "3", "4", "5"),
                       labels=c("Group 1", "Group 2", "Group 3", "Group 4", "Group 5")) +
  ggtitle("Movie segmentation", subtitle = "K-means Clustering")

```

### Conclusion

Speaking about our groups, we can interpret this result such as;

 * Group 1: Movies with medium  budget and medium gross
 * Group 2: Movies with low budget and high gross
 * Group 3: Movies with low budget and low gross
 * Group 4: Movies with medium budget and low gross
 * Group 5: Movies with low to medium budget and low gross
 
 
## Referances

 * https://www.datacamp.com/
 * Clustering Analysis in R in R Youtube by Hefin Rhys
 * kaggle
 * https://stackoverflow.com/
 * https://statsandr.com/blog/clustering-analysis-k-means-and-hierarchical-clustering-by-hand-and-in-r/
